---
title: "Homework2"
author: Phoebe Mo
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset

```{r}
trashwheel_df =
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017
```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation
```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information of the Mr. Trashwheel trash collector in Baltimore, Maryland. The trash collector will record how much trash it collects and stores in a dumpster. The dataset provides information of the year, month, and specific kinds of trash collected. There are totally `r nrow(trashwheel_df)` rows in the final dataset. There are additional data sheets that include month precipitation data. The total precipitation in 2018 is `r sum(pull(precip_2018, total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trashwheel_df, year == 2017), sports_balls))`.

## Problem 2

Read the NYC transit dataset

```{r}
subway_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude,
                   route1:route11, entry, entrance_type, vending, ada) %>%
  mutate(
    entry = recode(entry, YES = TRUE, NO = FALSE)
  )
```

This dataset contains information about the NYC transit subway stations.
In the dataset, information including lines, routes, stations and their locations are introduced. I have cleaned the names of each variable, selected major variables we want to look deeper, and changed the character input of entry into logical. I think the dataset is not yet tidy since there are 11 routes that are arranged in an untidy way.

There are `r unique(subway_df) %>% nrow` distinct stations in the dataset, and there are `r filter(subway_df, ada== TRUE) %>% nrow` stations are ADA compliant. The proportion of station entrances/exit without vending allow entrance is: `r unique(subway_df) %>% filter(vending=="NO", entry==TRUE) %>% nrow / unique(subway_df) %>% filter(vending=="NO") %>% nrow`

Now reformat the data so that route number and route name are distinct variables

```{r}
subway_df =
  mutate(subway_df, route8 = as.character(route8), route9 =   as.character(route9), route10 = as.character(route10), route11 = as.character(route11))

subway_tidy_df =
  pivot_longer(
    subway_df,
    route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  )
```

The are `r filter(subway_tidy_df, route_number=="A") %>% unique %>% nrow` distinct stations serve the A train. Of the stations that serve the A train, there are `r filter(subway_tidy_df, route_number=="A") %>% unique %>% filter(ada==TRUE) %>% nrow` of them are ADA compliant.

