---
title: "Homework2"
author: Phoebe Mo
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset

```{r}
trashwheel_df =
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017
```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation
```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information of the Mr. Trashwheel trash collector in Baltimore,
Maryland. The trash collector will record how much trash it collects and stores in a
dumpster. The dataset provides information of the year, month, and specific kinds of
trash collected. There are totally `r nrow(trashwheel_df)` rows in the final dataset. There are additional data sheets that include month precipitation data. The total
precipitation in 2018 is `r sum(pull(precip_2018, total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trashwheel_df, year == 2017), sports_balls))`.