---
title: "Homework2"
author: Phoebe Mo
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset

```{r}
trashwheel_df =
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017
```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation
```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017) %>%
  left_join(month_df, by = "month")
```

This dataset contains information of the Mr. Trashwheel trash collector in Baltimore, Maryland. The trash collector will record how much trash it collects and stores in a dumpster. The dataset provides information of the year, month, and specific kinds of trash collected. There are totally `r nrow(trashwheel_df)` rows in the final dataset. There are additional data sheets that include month precipitation data. The total precipitation in 2018 is `r sum(pull(precip_2018, total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trashwheel_df, year == 2017), sports_balls))`.

## Problem 2

Read the NYC transit dataset

```{r}
subway_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude,
                   route1:route11, entry, entrance_type, vending, ada) %>%
  mutate(
    entry = recode(entry, YES = TRUE, NO = FALSE)
  )
```

This dataset contains information about the NYC transit subway stations.
In the dataset, information including lines, routes, stations and their locations are introduced. There are totally `r nrow(subway_df)` rows in the final dataset. I have cleaned the names of each variable, selected major variables we want to look deeper, and changed the character input of entry into logical. I think the dataset is not yet tidy since there are 11 routes that are arranged in an untidy way.

There are `r distinct(subway_df, station_name, line, .keep_all=T) %>% nrow` distinct stations in the dataset, and there are `r filter(subway_df, ada==TRUE) %>% distinct(station_name, line) %>% nrow` stations are ADA compliant. The proportion of station entrances/exit without vending allow entrance is: `r filter(subway_df, vending=="NO", entry==TRUE) %>% nrow / filter(subway_df, vending=="NO") %>% nrow`

Now reformat the data so that route number and route name are distinct variables

```{r}
subway_df =
  mutate(subway_df, route8 = as.character(route8), route9 =   as.character(route9), route10 = as.character(route10), route11 = as.character(route11))

subway_tidy_df =
  pivot_longer(
    subway_df,
    route1:route11,
    names_to = "route_name",
    values_to = "route_number"
  )
```

The are `r filter(subway_tidy_df, route_number=="A") %>% distinct(station_name, line) %>% nrow` distinct stations serve the A train. Of the stations that serve the A train, there are `r filter(subway_tidy_df, route_number=="A", ada==TRUE) %>% distinct(station_name, line) %>% nrow` of them are ADA compliant.

## Problem 3

First, clean and manipulate data in pols-month.csv

```{r}
pols_month_df =
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "[-]") %>%
  mutate(month = as.numeric(month)) %>%
  left_join(month_df, by = "month") %>%
  select(-c(month, day)) %>%
  select(year, month_name, everything()) %>%
  rename(month = month_name) %>%
  mutate(
    president = prez_gop - prez_dem,
    president = ifelse(president == -1, "dem", "gop")
  ) %>%
  select(-c(prez_gop, prez_dem))
```

Second, clean and manipulate data in snp.csv

```{r}
snp_df =
  read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), sep = "[/]") %>%
  mutate(month = as.numeric(month)) %>%
  left_join(month_df, by = "month") %>%
  select(-c(month,day)) %>%
  select(year, month_name, everything()) %>%
  rename(month = month_name)
```

Third, clean and manipulate data in unemployment.csv
```{r}
unemployment_df =
  read_csv("./data/unemployment.csv")

colnames(unemployment_df) = c("year", month.name)

unemployment_df =
  pivot_longer(
    unemployment_df,
    January:December,
    names_to = "month",
    values_to = "unemployment"
  ) %>%
  mutate(year = as.character(year))
```

Join the datasets by merging snp into pols and merge "unemployment" into result
```{r}
bind_df =
  left_join(pols_month_df, snp_df, by = c("year", "month"))

bind_df = left_join(bind_df, unemployment_df, by = c("year", "month"))
```

The pols_month_df dataset contains information about the number of national politicians
who are either democratic or republican at a given time. The snp_df dataset contains
information of the representative measure of stock market (S&P stock index) at a given
time. The unemployment_df provides information of the unemployment rate at a given
time. The bind_df, which is a dataset combining information from these three datasets,
uses years and months as keys to provide the number of national politicians(dem or gop), S&P stock index, and unemployment rates in any given year and month.

There are totally `r nrow(bind_df)` rows and `r ncol(bind_df)` columns in the final
dataset, and variables include `r names(bind_df)`. The range of years is `r min(pull(bind_df, year))` to `r max(pull(bind_df, year))`. The average S&P stock index in these available years is `r mean(pull(bind_df, close), na.rm = T)` and the average unemployment rate in these available years is `r mean(pull(bind_df, unemployment), na.rm = T)` precent.
